### 2021-11-25 documentos/generators.tex
Agregada potencial figura para ver la simetría en la a y PCEGs de 1 qubit
Please no changes this time.

There exists a subset of PCE channels that generate 
all PCEs; the nature of these generators may be studied, as we
shall see, with the properties of the aforementioned vector space.
%\janote{Este párrafo lo voy a retocar. El enunciado de arriba es el primer enunciado
%del nuevo párrafo. Ojo que hay que introducir la notación $\mcG$.}
By standard theorems of linear algebra, any subspace $W$, see \Eref{eq:19}, 
can be extended to a maximal non-trivial subspace of dimension $2N-1$ by
adjoining appropriate additional basis elements. This can be done in
different ways. We therefore arrive to the set of maximal extensions of $W$,
where every maximal subspace corresponds to a PCE quantum channel that
preserves half of the Pauli components. The intersection of all the elements of
this set reduces to $W$ itself,
and since intersection of subspaces translates to composition of PCEs,
this implies that all PCEs \cpnote{damm la identidad} can be obtained as
compositions the PCEs corresponding to maximally non-trivial subspaces, 
which we call generators. 
\cpnote{Aca voy}
Consider \Fref{fig:examplesPCE}; the composition of any two 2-qubits PCE
generators 
in (c), (d), or (e) is represented by taking the intersection between them both
and the three different possible compositions result in PCE channel in (b).
The existence of the PCE generators $\mcG$ provide the program of 
constructing all PCE channels a tool that avoids the usage of a 
high-computational-cost method. In a moment, we will discuss how to find the 
generators $\mcG$ from matrix $A$, given a symmetry that is obeyed both by the generators 
$\mcG$ and matrix $A$ in \Eref{eq:san}.
%For the case of 1 qubit, the existence of PCE generators $\pceg$ is  
%verified by Siudzińska~\cite{siudzinska2019geometry}.
%\janote{Esto puede
%ser un párrafo, o puede ser un comentario final dentro del párrafo antrerior.}

PCE generators may be identified with a unique vector index
$\vec\alpha$, and thus be denoted as $\pceg$.
% El numero de generadores concide con esa vaina. 
%Let us introduce a convention for the following discussion for 
%the indices in $\vec\alpha$; the number 0 to 3 identifies
Recall that indices $0,1,2,3$ in $\vec\alpha$ correspond to, respectively,
the identity map and the 3 1-qubit PCE quantum channels that project the 
Bloch sphere to the rectangular axes $x$, $y$ and $z$. 
Therefore we will correspond every PCE generator to a collection of numbers 
that tells what its local-action is on every individual qubit in the system.
All of this collections of numbers happen to be all vector indices $\{\valpha\}$.
For example, the PCE generator represented in \Fref{fig:examplesPCE}(c),
whose local-action on the first qubit (first column) is to project the corresponding local Bloch sphere 
to the $x$ axis and on the second qubit (first row) to map the corresponding
local Bloch sphere to itself, may be characterized with the vector index (1,0). 
Then, it is clear that in general one can associate a unique $\vec\alpha$ to 
every PCE generator and so denote them as $\pceg$.

PCE generators $\pceg$ obey a symmetry
that defines uniquely every PCE generator. The coefficients  $\{\taus\}$ of every 
generator $\pceg$ arranged in a hypercubic $N$-dimensional array 
%of dimensions $\underbrace{4\times 4\times\ldots\times 4}_{N}$ 
obey a symmetry with respect to the $N$ hyperplanes of dimension $N-1$
\janote{según wikipedia, por definición, los hiperplanos son de dimensión 
$N-1$. Así que podría quitarse la aclaración de la dimensión. Lo dejo así y 
espero retroalimentación} 
that separate the array into two half arrays \janote{No estoy tan satisfecho con la redacción de este enunciado}.
See, in \Fref{fig:examplesPCE}, that generators $\mcG_{(1,3)}$, $\mcG_{(3,2)}$
and $\mcG_{(2,2)}$ represented in (c), (d) and (e), respectively,
are either symmetric or anti-symmetric with respect to the lines that cut vertically 
and horizontally in half the diagrams. In fact, this symmetry may be used to
construct all generators $\pceg$ as follows. Start out with $\tau_{0,\ldots,0}=1$
and begin to ``unfold'' the array of all $\pceg$'s coefficients $\taus$ by deciding
whether $\tau_{0,\ldots,0,1}$ be set to 1 or 0 (symmetric or anty-simmetric with 
respect to $\tau_{0,\ldots,0}$), then choose the values of $\{\tau_{0,\ldots,0,2}, 
\tau_{0,\ldots,0,3}\}$ to be symmetric or anti-symmetric with respect to 
$\{\tau_{0,\ldots,0},\tau_{0,\ldots,0,1}\}$; you need to keep this procedure 
for the remaining subsets with $2l$ coefficients $\taus$, with $l$ the cardinality of 
the previous subset, by setting the coefficients $\taus=0,1$ in order to be symmetric or
anti-symmetric to the previous subset. For example, let us illustrate the 
procedure to construct $\mcG_{(3,2)}$ represented in \Fref{fig:examplesPCE}(d).
One starts out choosing the value of $\tau_{(0,1)}$ to be anti-symmetric to $\tau_{(0,0)}$, then one chooses 
$\qty{\tau_{(0,2)},\tau_{(0,3)}}$ symmetric to $\qty{\tau_{(0,0)},\tau_{(0,1)}}$,
$\qty{\tau_{(1,0)},\tau_{(1,1)},\tau_{(1,2)},\tau_{(1,3)}}$ anti-symmetric 
to $\qty{\tau_{(0,0)},\tau_{(0,1)},\tau_{(0,2)},\tau_{(0,3)}}$, and finally
the last eight components $\tau_{\vec\alpha}$ anti-symmetric to the first 
eight.

The matrix $\san$ [see \eref{eq:san}] encodes all the information
of PCE generators $\pceg$ and, therefore, of all PCE quantum channels. 
The rows and columns of matrix $A$ also obey the symmetry previously discussed 
for PCE generators $\pceg$. It suffices to check that rows and columns of matrix $a$
in \Eref{eq:1} obey the aforementioned symmetry when they are arranged in a hypercubic 
$N$-dimensional array of order $4$.
As a consequence of $a$'s elements being $\pm1$, the power tensor $a^{\otimes N}$ will hold the symmetry of the 
rows and columns of $A$. 
Therefore, this establishes a deep connection between the matrix $A$ and 
PCE quantum channels. It allows one to infer the PCE generators $\pceg$
by taking its corresponding coefficients $\{\taus\}$ as the rows or columns of $A$ 
after taking $-1\mapsto0$. In \appref{sec:2_qubits} we illustrate this program 
for 2-qubits PCE channels.
In short, it is sufficient to have the matrix $A$ to construct 
all PCE quantum channels.
%\janote{Aquí
%tendrá que ir una prueba que aún se está cocinando}

>>>>

There exists a subset of PCE channels that generate 
all PCEs; the nature of these generators may be studied, as we
shall see, with the properties of the aforementioned vector space.
%\janote{Este párrafo lo voy a retocar. El enunciado de arriba es el primer enunciado
%del nuevo párrafo. Ojo que hay que introducir la notación $\mcG$.}
By standard theorems of linear algebra, any proper subspace $W$, see \Eref{eq:19}, 
can be extended to a maximal non-trivial subspace of dimension $2N-1$ by
adjoining appropriate additional basis elements. This can be done in
different ways. We therefore arrive to the set of maximal extensions of $W$,
where every maximal subspace corresponds to a PCE quantum channel that
preserves half of the Pauli components. The intersection of all the elements of
this set reduces to $W$ itself,
and since intersection of subspaces translates to composition of PCEs,
this implies that all PCEs can be obtained as
compositions the PCEs corresponding to maximally non-trivial subspaces, plus
the identity map. In other words, the set of PCEs that preserve half 
of the components plus the identity map, are a generator set for 
all PCE channels.
% and the complete vector space (corresponding to the identity channel),
% which we call generators. 
% \cpnote{Aca tenemos un semigrupo y no se si incluir a la identidad como 
% generador. De alguna manera es excepcional. Es problematico incluirla y es 
% problematico no incluirla.}
Consider \Fref{fig:examplesPCE}; subfigures (c), (d) and (e) represent nontrivial 
generators and the composition of any two of them yield the PCE channel
corresponding to (b). 
% 
% 
% the composition of any two 2-qubits PCE
% generators 
% in (c), (d), or (e) is equivalent to taking the intersection between them both
% and the three different possible compositions result in PCE channel in (b).
\cpnote{Aca se me hace que hay un cambio brusco de tema. Ver que onda
con Jose Alfredo}
The existence of the PCE generators $\mcG$ provide the program of 
constructing all PCE channels a tool that avoids the usage of a 
high-computational-cost method. In a moment, we will discuss how to find the 
generators $\mcG$ from matrix $A$, given a symmetry that is obeyed both by the generators 
$\mcG$ and matrix $A$ in \Eref{eq:san}\cpnote{Como que para que ahora se
menciona entonces?}.
%For the case of 1 qubit, the existence of PCE generators $\pceg$ is  
%verified by Siudzińska~\cite{siudzinska2019geometry}.
%\janote{Esto puede
%ser un párrafo, o puede ser un comentario final dentro del párrafo antrerior.}

\cpnote{Arreglar}
PCE generators may be identified with a unique vector index
$\vec\alpha$, and thus be denoted as $\pceg$.
% El numero de generadores concide con esa vaina. 
%Let us introduce a convention for the following discussion for 
%the indices in $\vec\alpha$; the number 0 to 3 identifies
Recall that indices $0,1,2,3$ in $\vec\alpha$ correspond to, respectively,
the identity map and the 3 1-qubit PCE quantum channels that project the 
Bloch sphere to the rectangular axes $x$, $y$ and $z$. 
Therefore we will correspond every PCE generator to a collection of numbers 
that tells what its local-action is on every individual qubit in the system.
All of this collections of numbers happen to be all vector indices $\{\valpha\}$.
For example, the PCE generator represented in \Fref{fig:examplesPCE}(c),
whose local-action on the first qubit (first column) is to project the corresponding local Bloch sphere 
to the $x$ axis and on the second qubit (first row) to map the corresponding
local Bloch sphere to itself, may be characterized with the vector index (1,0). 
Then, it is clear that in general one can associate a unique $\vec\alpha$ to 
every PCE generator and so denote them as $\pceg$.

\cpnote{Mucha lora y no se prueba nada}
PCE generators $\pceg$ obey a symmetry
that defines uniquely every PCE generator. The coefficients  $\{\taus\}$ of every 
generator $\pceg$ arranged in a hypercubic $N$-dimensional array 
%of dimensions $\underbrace{4\times 4\times\ldots\times 4}_{N}$ 
obey a symmetry with respect to the $N$ hyperplanes of dimension $N-1$
\janote{según wikipedia, por definición, los hiperplanos son de dimensión 
$N-1$. Así que podría quitarse la aclaración de la dimensión. Lo dejo así y 
espero retroalimentación} 
that separate the array into two half arrays \janote{No estoy tan satisfecho con la redacción de este enunciado}.
See, in \Fref{fig:examplesPCE}, that generators $\mcG_{(1,3)}$, $\mcG_{(3,2)}$
and $\mcG_{(2,2)}$ represented in (c), (d) and (e), respectively,
are either symmetric or anti-symmetric with respect to the lines that cut vertically 
and horizontally in half the diagrams. In fact, this symmetry may be used to
construct all generators $\pceg$ as follows. Start out with $\tau_{0,\ldots,0}=1$
and begin to ``unfold'' the array of all $\pceg$'s coefficients $\taus$ by deciding
whether $\tau_{0,\ldots,0,1}$ be set to 1 or 0 (symmetric or anty-simmetric with 
respect to $\tau_{0,\ldots,0}$), then choose the values of $\{\tau_{0,\ldots,0,2}, 
\tau_{0,\ldots,0,3}\}$ to be symmetric or anti-symmetric with respect to 
$\{\tau_{0,\ldots,0},\tau_{0,\ldots,0,1}\}$; you need to keep this procedure 
for the remaining subsets with $2l$ coefficients $\taus$, with $l$ the cardinality of 
the previous subset, by setting the coefficients $\taus=0,1$ in order to be symmetric or
anti-symmetric to the previous subset. For example, let us illustrate the 
procedure to construct $\mcG_{(3,2)}$ represented in \Fref{fig:examplesPCE}(d).
One starts out choosing the value of $\tau_{(0,1)}$ to be anti-symmetric to $\tau_{(0,0)}$, then one chooses 
$\qty{\tau_{(0,2)},\tau_{(0,3)}}$ symmetric to $\qty{\tau_{(0,0)},\tau_{(0,1)}}$,
$\qty{\tau_{(1,0)},\tau_{(1,1)},\tau_{(1,2)},\tau_{(1,3)}}$ anti-symmetric 
to $\qty{\tau_{(0,0)},\tau_{(0,1)},\tau_{(0,2)},\tau_{(0,3)}}$, and finally
the last eight components $\tau_{\vec\alpha}$ anti-symmetric to the first 
eight.

\begin{figure} % {{{
\centering
\includegraphics[width=\columnwidth]{oneQ_PCEGs-crop}
\caption{\janote{Propuesta de figura para mostrar que generadores 
y la a obedecen las mismas simetrías.}}
\end{figure} % }}}

\cpnote{Mucha lora y no se prueba nada, again}
The matrix $\san$ [see \eref{eq:san}] encodes all the information
of PCE generators $\pceg$ and, therefore, of all PCE quantum channels. 
The rows and columns of matrix $A$ also obey the symmetry previously discussed 
for PCE generators $\pceg$. It suffices to check that rows and columns of matrix $a$
in \Eref{eq:1} obey the aforementioned symmetry when they are arranged in a hypercubic 
$N$-dimensional array of order $4$.
As a consequence of $a$'s elements being $\pm1$, the power tensor $a^{\otimes N}$ will hold the symmetry of the 
rows and columns of $A$. 
Therefore, this establishes a deep connection between the matrix $A$ and 
PCE quantum channels. It allows one to infer the PCE generators $\pceg$
by taking its corresponding coefficients $\{\taus\}$ as the rows or columns of $A$ 
after taking $-1\mapsto0$. In \appref{sec:2_qubits} we illustrate this program 
for 2-qubits PCE channels.
In short, it is sufficient to have the matrix $A$ to construct 
all PCE quantum channels.
%\janote{Aquí
%tendrá que ir una prueba que aún se está cocinando}

\cpnote{Sienot que no tiene flujo. Quizá se le pueda dar flujo usando el final 
del primer parrafo}
